---
title: "Data 607 - Week4 - Project1"
author: "Glen Dale Davis"
date: "2023-02-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the Required Packages

Below, I load the packages required for tidying up and viewing the data.

```{r packages, warning = FALSE, message = FALSE}
library(tidyverse)
library(DT)

```

## Load and Clean Up the Raw Chess Tournament Info Data

Below I load the chess tournament info .txt file, replace all -...- lines with $*$, define $*$ as a comment character so those lines can then be skipped completely since they contain no useful info, and define the field separator as $|$ so a (still raw, but better) data frame can be created from the useful tournament info. I also trim whitespace from the data fields.

```{r raw_data, warning = FALSE}

my_url <- "https://raw.githubusercontent.com/geedoubledee/data607_week4/main/tournamentinfo.txt"
raw_chess_data <- readLines(con = my_url, warn = FALSE)
raw_chess_data <- str_replace_all(raw_chess_data, "^--*-$", "*")
writeLines(raw_chess_data, con = "tournamentinfo_altered.txt")
my_url2 <- "tournamentinfo_altered.txt"
altered_chess_data <- read.table(file = my_url2, header = TRUE, sep = "|",
                                 stringsAsFactors = FALSE, fill = TRUE,
                                 strip.white = TRUE,
                                 blank.lines.skip = TRUE,
                                 comment.char = "*")
as_tibble(altered_chess_data)

```

## Capture Data Fields from the Even and Odd Rows of the Altered Chess Tournament Info Data Frame as Column Vectors

Since there is different information that can be captured from the even rows in my altered data frame than from the odd rows, I loop through them separately below to capture the information I want. Player name, total number of points, and the opponents that player played all come from the even rows. Player state and player pre-rating come from the odd rows.

```{r altered_df_even_and_odd_row_fields, warning = FALSE}
even_rows <- seq(2, 128, 2)
player_names <- c()
total_nums_points <- c()
opponents_sets <- list()
count <- 0

for (i in even_rows){
    count <- count + 1
    player_name <- altered_chess_data[[i, 2]]
    player_names <- append(player_names, player_name)
    total_num_points <- altered_chess_data[[i, 3]]
    total_nums_points <- append(total_nums_points, total_num_points)
    opponents_set <- c()
    for (j in 4:10){
        opponent <- altered_chess_data[[i, j]]
        opponent <- as.integer(str_replace(opponent, ". ", ""))
        opponents_set <- append(opponents_set, opponent)
    }
    opponents_sets[[count]] <- opponents_set
}

odd_rows <- seq(3, 129, 2)
player_states <- c()
player_pre_ratings <- c()

for (i in odd_rows){
    player_state <- altered_chess_data[[i, 1]]
    player_states <- append(player_states, player_state)
    player_pre_rating <- altered_chess_data[[i, 2]]
    player_pre_rating <- str_replace(player_pre_rating, ".*R: ", "")
    player_pre_rating <- str_replace(player_pre_rating, "P[0-9]*", "")
    player_pre_rating <- str_replace(player_pre_rating, "->.*", "")
    player_pre_rating <- as.integer(trimws(player_pre_rating))
    player_pre_ratings <- append(player_pre_ratings, player_pre_rating)
}

```

## Calculate The Average Pre-Chess Rating of the Opponents for Each Player and Store Info in Column Vector

Below I use the sets of opponents I captured earlier for each player to determine the average pre-chess rating of that player's opponents. This is the final column vector needed to create the proper chess tournament results data frame I've been working toward. 

```{r avg_pre_chess_rating_vector}
avg_pre_chess_ratings_of_opponents <- c()

for (i in 1:length(opponents_sets)){
    set <- opponents_sets[[i]]
    ratings <- c()
    for (j in set){
        if (!is.na(j)){
            rating <- player_pre_ratings[j]
            ratings <- append(ratings, rating)
        }
    }
    avg_pre_chess_rating_of_opponents <- round(mean(ratings))
    avg_pre_chess_ratings_of_opponents <-
        append(avg_pre_chess_ratings_of_opponents,
               avg_pre_chess_rating_of_opponents)
}

```

## Combine the Column Vectors into a Final Chess Tournament Results Data Frame

Now the column vectors can be combined into the final chess tournament results data frame, from which I create a data table and display 15 rows at a time. 

```{r final_chess_tournament_df}

chess_df <- as.data.frame(
    cbind(player_names, player_states, total_nums_points,
          player_pre_ratings, avg_pre_chess_ratings_of_opponents))
cols <- c("Player_Name", "Player_State", "Total_Number_of_Points",
          "Player_Pre_Rating", "Average_Pre_Chess_Rating_of_Opponents")
colnames(chess_df) <- cols
chess_df$Player_Pre_Rating <- as.integer(chess_df$Player_Pre_Rating)
chess_df$Average_Pre_Chess_Rating_of_Opponents <- as.integer(
    chess_df$Average_Pre_Chess_Rating_of_Opponents)
datatable(chess_df, options = list(pageLength = 15))

```

## Export the Final Chess Tournament Results Data Frame

Now the final chess tournament results data frame can be exported to .csv.


```{r export_csv}

write.csv(chess_df, "chess_tournament_results.csv")

```